---
title: "Assignment1"
author: "Stephen Jones"
date: "January 31, 2019"
output: html_document
---
# <span style="font-size:10px;">Choose from the tabs below to view code and links to files on github.</span> {.tabset}

## Assignment 1


```{r message=FALSE, warning=FALSE}

#clean the workspace.
rm(list=ls())
getwd()
#need data.table for fread.
suppressWarnings(library(data.table))
```

```{r eval=FALSE}
#download .data file from url, write to local drive, then upload to github
mushdata<-fread("https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data")
write.csv(mushdata,'agaricus-lepiota.csv')
```

```{r message=FALSE, warning=FALSE}
#download data from github
library(RCurl)
mushdata<-read.csv(text=getURL("https://raw.githubusercontent.com/sigmasigmaiota/mushroom/master/agaricus-lepiota.csv"))

#examine data
summary(mushdata)
head(mushdata)

#remove first column which serves as a redundant row counter.
mushdata$X<-NULL
```
#####Using the codebook as a reference, test columns for unique values to verify; do the variables match the codebook?
```{r message=FALSE, warning=FALSE}
sapply(mushdata, function(x) unique(x))
```
#####I'd like to examine the data describing traits observable to pedestrians or hikers, renaming those variables which have been verified with the codebook.
```{r message=FALSE, warning=FALSE}

colnames(mushdata)[1]<-"Edible"
colnames(mushdata)[2]<-"CapShape"
colnames(mushdata)[3]<-"CapSurface"
colnames(mushdata)[4]<-"CapColor"
colnames(mushdata)[6]<-"Odor"
colnames(mushdata)[22]<-"Population"
colnames(mushdata)[23]<-"Habitat"
```
#####Check missing values before recoding. There are 0.
```{r message=FALSE, warning=FALSE}
#Check missing values
sapply(mushdata, function(x) sum(is.na(x)))
```
#####Recode using recode command from "car" package.
```{r message=FALSE, warning=FALSE}
library(car)
mushdata$Edible<-recode(mushdata$Edible,
                        "'e'='edible';
                        'p'='poisonous'")
mushdata$CapShape<-recode(mushdata$CapShape,
                        "'b'='bell';
                        'c'='conical';
                        'x'='convex';
                        'f'='flat';
                        'k'='knobbed';
                        's'='sunken'")

mushdata$CapSurface<-recode(mushdata$CapSurface,
                        "'f'='fibrous';
                        'g'='grooves';
                        'y'='scaly';
                        's'='smooth'")

mushdata$CapColor<-recode(mushdata$CapColor,
                        "'n'='brown';
                        'b'='buff';
                        'c'='cinnamon';
                        'g'='gray';
                        'r'='green';
                        'p'='pink';
                        'u'='purple';
                        'e'='red';
                        'w'='white';
                        'y'='yellow'")
mushdata$Odor<-recode(mushdata$Odor,
                        "'a'='almond';
                        'l'='anise';
                        'c'='creosote';
                        'y'='fishy';
                        'f'='foul';
                        'm'='musty';
                        'n'='none';
                        'p'='pungent';
                        's'='spicy'")
mushdata$Population<-recode(mushdata$Population,
                        "'a'='abundant';
                        'c'='clustered';
                        'n'='numerous';
                        's'='scattered';
                        'v'='several';
                        'y'='solitary'")
mushdata$Habitat<-recode(mushdata$Habitat,
                         "'g'='grasses';
                         'l'='leaves';
                         'm'='meadows';
                         'p'='paths';
                         'u'='urban';
                         'w'='waste';
                         'd'='woods'")

```
#####Check missing values after recoding.
```{r message=FALSE, warning=FALSE}
sapply(mushdata, function(x) sum(is.na(x)))

```
#####Create subset with selected variables.
```{r message=FALSE, warning=FALSE}
mushdata.sub<-mushdata[,c("Edible","CapShape","CapSurface","CapColor","Odor","Population","Habitat")]
```
#####Create additional numeric variables to prepare for analysis.
```{r message=FALSE, warning = FALSE}
mushdata.sub$EdibleN<-as.numeric(recode(mushdata$Edible,
                        "'edible'=1;
                        'poisonous'=0"))
mushdata.sub$CapShapeN<-as.numeric(recode(mushdata$CapShape,
                        "'bell'=1;
                        'conical'=2;
                        'convex'=3;
                        'flat'=4;
                        'knobbed'=5;
                        'sunken'=6"))

mushdata.sub$CapSurfaceN<-as.numeric(recode(mushdata$CapSurface,
                        "'fibrous'=1;
                        'grooves'=2;
                        'scaly'=3;
                        'smooth'=4"))

mushdata.sub$CapColorN<-as.numeric(recode(mushdata$CapColor,
                        "'brown'=1;
                        'buff'=2;
                        'cinnamon'=3;
                        'gray'=4;
                        'green'=5;
                        'pink'=6;
                        'purple'=7;
                        'red'=8;
                        'white'=9;
                        'yellow'=10"))
mushdata.sub$OdorN<-as.numeric(recode(mushdata$Odor,
                        "'almond'=1;
                        'anise'=2;
                        'creosote'=3;
                        'fishy'=4;
                        'foul'=5;
                        'musty'=6;
                        'none'=7;
                        'pungent'=8;
                        'spicy'=9"))
mushdata.sub$PopulationN<-as.numeric(recode(mushdata$Population,
                        "'abundant'=1;
                        'clustered'=2;
                        'numerous'=3;
                        'scattered'=4;
                        'several'=5;
                        'solitary'=6"))
mushdata.sub$HabitatN<-as.numeric(recode(mushdata$Habitat,
                         "'grasses'=1;
                         'leaves'=2;
                         'meadows'=3;
                         'paths'=4;
                         'urban'=5;
                         'waste'=6;
                         'woods'=7"))
```
#####Verify numeric recoded variables.
```{r message=FALSE, warning = FALSE}
sapply(mushdata.sub[,c("EdibleN","CapShapeN","CapSurfaceN","CapColorN","OdorN","PopulationN","HabitatN")], function(x) is.numeric(x))
```

#####Split the dataframe, in preparation to compare edible and poisonous plots.
```{r message=FALSE, warning = FALSE}

Edible.Pois<-split(mushdata.sub[2:14],mushdata$Edible)
pois<-as.data.frame(Edible.Pois$poisonous)
edib<-as.data.frame(Edible.Pois$edible)

```

## Links
Link to repository file:
https://github.com/sigmasigmaiota/mushroom/blob/master/Assignment1_StephenJones.Rmd

Direct link to this .Rmd:
https://raw.githubusercontent.com/sigmasigmaiota/mushroom/master/Assignment1_StephenJones.Rmd

