---
title: "Assignment1"
author: "Stephen Jones"
date: "January 31, 2019"
output:
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
---
# <span style="font-size:10px;">Choose from the tabs below to view code and links to files on github.</span> {.tabset}

## Assignment 1


```{r message=FALSE, warning=FALSE}
#clean the workspace.
rm(list=ls())
getwd()
#need data.table for fread.
suppressWarnings(library(data.table))
```

```{r eval=FALSE}
#download .data file from url, write to local drive, then upload to github
mushdata<-fread("https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data")
write.csv(mushdata,'agaricus-lepiota.csv')
```

```{r message=FALSE, warning=FALSE}
#download data from github
library(RCurl)
mushdata<-read.csv(text=getURL("https://raw.githubusercontent.com/sigmasigmaiota/mushroom/master/agaricus-lepiota.csv"))

#examine data
summary(mushdata)
head(mushdata)

#remove first column which serves as a redundant row counter.
mushdata$X<-NULL
```
#####Using the codebook as a reference, test columns for unique values to verify; do the variables match the codebook?
```{r message=FALSE, warning=FALSE}
sapply(mushdata, function(x) unique(x))
```
#####I'd like to examine the data describing traits observable to pedestrians or hikers, renaming those variables which have been verified with the codebook.
```{r message=FALSE, warning=FALSE}

colnames(mushdata)[1]<-"Edible"
colnames(mushdata)[2]<-"CapShape"
colnames(mushdata)[3]<-"CapSurface"
colnames(mushdata)[4]<-"CapColor"
colnames(mushdata)[6]<-"Odor"
colnames(mushdata)[22]<-"Population"
colnames(mushdata)[23]<-"Habitat"
```
#####Check missing values before recoding. There are 0.
```{r message=FALSE, warning=FALSE}
#Check missing values
sapply(mushdata, function(x) sum(is.na(x)))
```
#####Recode using recode command from "car" package.
```{r message=FALSE, warning=FALSE}
library(car)
mushdata$Edible<-recode(mushdata$Edible,
                        "'e'='edible';
                        'p'='poisonous'")
mushdata$CapShape<-recode(mushdata$CapShape,
                        "'b'='bell';
                        'c'='conical';
                        'x'='convex';
                        'f'='flat';
                        'k'='knobbed';
                        's'='sunken'")

mushdata$CapSurface<-recode(mushdata$CapSurface,
                        "'f'='fibrous';
                        'g'='grooves';
                        'y'='scaly';
                        's'='smooth'")

mushdata$CapColor<-recode(mushdata$CapColor,
                        "'n'='brown';
                        'b'='buff';
                        'c'='cinnamon';
                        'g'='gray';
                        'r'='green';
                        'p'='pink';
                        'u'='purple';
                        'e'='red';
                        'w'='white';
                        'y'='yellow'")
mushdata$Odor<-recode(mushdata$Odor,
                        "'a'='almond';
                        'l'='anise';
                        'c'='creosote';
                        'y'='fishy';
                        'f'='foul';
                        'm'='musty';
                        'n'='none';
                        'p'='pungent';
                        's'='spicy'")
mushdata$Population<-recode(mushdata$Population,
                        "'a'='abundant';
                        'c'='clustered';
                        'n'='numerous';
                        's'='scattered';
                        'v'='several';
                        'y'='solitary'")
mushdata$Habitat<-recode(mushdata$Habitat,
                         "'g'='grasses';
                         'l'='leaves';
                         'm'='meadows';
                         'p'='paths';
                         'u'='urban';
                         'w'='waste';
                         'd'='woods'")

```
#####Check missing values after recoding.
```{r message=FALSE, warning=FALSE}
sapply(mushdata, function(x) sum(is.na(x)))

```
#####Create subset with selected variables.
```{r message=FALSE, warning=FALSE}
mushdata.sub<-mushdata[,c("Edible","CapShape","CapSurface","CapColor","Odor","Population","Habitat")]
```

#####Split the dataframe, in preparation to compare edible and poisonous plots.
```{r message=FALSE, warning = FALSE}

Edible.Pois<-split(mushdata.sub[2:7],mushdata$Edible)
pois<-as.data.frame(Edible.Pois$poisonous)
edib<-as.data.frame(Edible.Pois$edible)

CS.p<-table(pois$CapShape)
CSur.p<-table(pois$CapSurface)
CC.p<-table(pois$CapColor)
OD.p<-table(pois$Odor)
POP.p<-table(pois$Population)
HB.p<-table(pois$Habitat)
CS.e<-table(edib$CapShape)
CSur.e<-table(edib$CapSurface)
CC.e<-table(edib$CapColor)
OD.e<-table(edib$Odor)
POP.e<-table(edib$Population)
HB.e<-table(edib$Habitat)

Prop.CapShape<-merge(as.data.frame(prop.table(CS.p)),as.data.frame(prop.table(CS.e)),by="Var1")
Prop.CapSurface<-merge(as.data.frame(prop.table(CSur.p)),as.data.frame(prop.table(CSur.e)),by="Var1")
Prop.CapColor<-merge(as.data.frame(prop.table(CC.p)),as.data.frame(prop.table(CC.e)),by="Var1")
Prop.Odor<-merge(as.data.frame(prop.table(OD.p)),as.data.frame(prop.table(OD.e)),by="Var1")
Prop.Population<-merge(as.data.frame(prop.table(POP.p)),as.data.frame(prop.table(POP.e)),by="Var1")
Prop.Habitat<-merge(as.data.frame(prop.table(HB.p)),as.data.frame(prop.table(HB.e)),by="Var1")

names(Prop.CapShape)<-c("Quality","Pois","Edib")
names(Prop.CapSurface)<-c("Quality","Pois","Edib")
names(Prop.CapColor)<-c("Quality","Pois","Edib")
names(Prop.Odor)<-c("Quality","Pois","Edib")
names(Prop.Population)<-c("Quality","Pois","Edib")
names(Prop.Habitat)<-c("Quality","Pois","Edib")
```

```{r message=FALSE, warning = FALSE}
library(ggplot2)

ggplot(Prop.CapShape) + 
  geom_point(aes(Quality,Edib),shape=69,fill="darkgreen", color="green", size=3,alpha=.8) + 
  geom_point(aes(Quality,Pois),shape=80,fill="darkred", color="red", size=3, alpha=.8) +
  theme_bw()+
  theme(panel.border = element_blank(),
        axis.line = element_line(colour = "black"))+
  ggtitle("Mushroom Cap Shape, Edible vs Poisonous",
          subtitle="https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data")+
  xlab("Quality")+
  ylab("Percent")
```







```{r message=FALSE, warning = FALSE}
ggplot(Prop.CapColor) + 
  geom_point(aes(Quality,Edib),shape=69,fill="darkgreen", color="green", size=3,alpha=.8) + 
  geom_point(aes(Quality,Pois),shape=80,fill="darkred", color="red", size=3, alpha=.8) +
  theme_bw()+
  theme(panel.border = element_blank(),
        axis.line = element_line(colour = "black"))+
  ggtitle("Mushroom Cap Color, Edible vs Poisonous",
          subtitle="https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data")+
  xlab("Quality")+
  ylab("Percent")
```

```{r message=FALSE, warning = FALSE}
ggplot(Prop.CapSurface) + 
  geom_point(aes(Quality,Edib),shape=69,fill="darkgreen", color="green", size=3,alpha=.8) + 
  geom_point(aes(Quality,Pois),shape=80,fill="darkred", color="red", size=3, alpha=.8) +
  theme_bw()+
  theme(panel.border = element_blank(),
        axis.line = element_line(colour = "black"))+
  ggtitle("Mushroom Cap Color, Edible vs Poisonous",
          subtitle="https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data")+
  xlab("Quality")+
  ylab("Percent")
```

```{r message=FALSE, warning = FALSE}
ggplot(Prop.Habitat) + 
  geom_point(aes(Quality,Edib),shape=69,fill="darkgreen", color="green", size=3,alpha=.8) + 
  geom_point(aes(Quality,Pois),shape=80,fill="darkred", color="red", size=3, alpha=.8) +
  theme_bw()+
  theme(panel.border = element_blank(),
        axis.line = element_line(colour = "black"))+
  ggtitle("Mushroom Cap Color, Edible vs Poisonous",
          subtitle="https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data")+
  xlab("Quality")+
  ylab("Percent")
```

```{r message=FALSE, warning = FALSE}
ggplot(Prop.Odor) + 
  geom_point(aes(Quality,Edib),shape=69,fill="darkgreen", color="green", size=3,alpha=.8) + 
  geom_point(aes(Quality,Pois),shape=80,fill="darkred", color="red", size=3, alpha=.8) +
  theme_bw()+
  theme(panel.border = element_blank(),
        axis.line = element_line(colour = "black"))+
  ggtitle("Mushroom Cap Color, Edible vs Poisonous",
          subtitle="https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data")+
  xlab("Quality")+
  ylab("Percent")
```

```{r message=FALSE, warning = FALSE}
ggplot(Prop.Population) + 
  geom_point(aes(Quality,Edib),shape=69,fill="darkgreen", color="green", size=3,alpha=.8) + 
  geom_point(aes(Quality,Pois),shape=80,fill="darkred", color="red", size=3, alpha=.8) +
  theme_bw()+
  theme(panel.border = element_blank(),
        axis.line = element_line(colour = "black"))+
  ggtitle("Mushroom Cap Color, Edible vs Poisonous",
          subtitle="https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data")+
  xlab("Quality")+
  ylab("Percent")
```

## Links
Link to repository .Rmd file:
https://github.com/sigmasigmaiota/mushroom/blob/master/Assignment1_StephenJones.Rmd

Direct link to this .Rmd:
https://raw.githubusercontent.com/sigmasigmaiota/mushroom/master/Assignment1_StephenJones.Rmd

